Upwelling variability across the Benguela upwelling systems

### 2019-10-07 SST_patterns.Rmd
updated

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
  theme(strip.text.x = element_text(size = 25))
p.wr3

@@

```{r}
anova_func <- function(df){
  sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration * distance, data = df)
return(sites_aov)
}

@@

SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * lubridate::year(date_start) * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```

>>>>

p.wr3 <- p.wr2 + facet_wrap(.~ site, ncol = 2, nrow = 2) +
  theme(strip.text.x = element_text(size = 25)) + theme(panel.spacing = unit(2, "lines"))
p.wr3

@@

```{r}
anova_func <- function(df){
  sites_aov <- aov(intensity_mean ~ site * duration * distance, data = df)
return(sites_aov)
}

@@

SACTN_anov <- sites_aov <- aov(intensity_mean ~ site * duration, data = SACTN_upwell_base)
summary(SACTN_anov)
```

### 2019-10-07 Write_up/Paper1_writeup.Rmd
updated

Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this). These temperatures have increased as a result of anthropogenically driven climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed). This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. Climate change impacting oceanic processes has major ecological consequences [@walther]. This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance. Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters [@hoegh]. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper. Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions. (RWS: There appear to be two or three separate main points being made in this paragraph. I would recomend you use what you have here and try and work these main points into other paragraphs)

@@

(RWS: This is a very thorough introduction section. Impotantly see that they all flow from one paragraph to the next. If it can be re-order a bit I think that it would be pretty close to a final version once the needed citations are inserted.)

@@

>>>>

Several studies have provided significant evidence suggesting that there has been an increase in ocean surface temperatures over the last few decades (Add references; RWS: You can cite the IPCC for this). These temperatures have increased as a result of anthropogenically driven climate change [@gruber]. Warming is shown to slowly move down into the water column thus increasing thermal stratification (RWS: citation needed). This thermal stratification restricts the depth at which water is upwelled and therfore limits the amount of nutrients brought up to the surface with a decrease in oxygen and pH levels [@chhak; @jacox2011; @jocox2016]. Climate change impacting oceanic processes has major ecological consequences [@walther]. This is demonstrated in several regions increases in sea surface temperatures have been proven to greatly influence marine species survival, reproduction and abundance. Climate change has also been shown to largely influence coastal species such as corals, muscles and oysters [@hoegh]. Given the importance of primary production, it was determined that the Benguela upwelling system, the eastern boundary upwelling system in South Africa, should be the focus of this paper. Furthermore, it should be understood that increases in sea surface temperatures bear considerable economic impacts as well [@murawski2000; @murawski1993] therefore understanding the impact and cause of climate change has merit not only in mitigation efforts in limiting environmental degradation and maintaining ecosystem functioning, but in its negative affects to local and broader economies as well as human populations dependent on these coastal regions.

@@

@@

## Statistical analyses

A series of ANOVA tests were used to compare the main effects of the chosen variables on a continuous variable. In this analysis the relationship between sites based on the mean intenisity of upwelling as a function of distance and duration were analysed; these analyses tested if significant
differences occurred between upwelling at different sites and distances from the coastline within each of the various datasets.

## Results

# Table

# Wind rose diagrams

```{r , echo= FALSE, message=FALSE, warning=FALSE, fig.cap="Wind rose diagram representing the most predominant wind direction for each of the sites. Each spoke is divided by colour into wind speed ranges. The radial length of each spoke around the circle is the time that the wind comes from that direction", fig.height=8, fig.width=12, fig.pos="H"}

knitr::include_graphics("../Figures/wind_roses.png")

```

Wind diagrams help visualise the patterns present at a particular site. Moving outward on the radial scale, the frequency associated with wind and waves coming from a particular direction increases. The predominant wind direction at Port Nolloth is 135o and at Sea Point it is 315o. Sea Point and Port Nollowth shows the fastest wind speed at a direction of 135o. The predominant wind direction for Lamberts Bay and Saldanha Bay occured at 195o, with the highest wind speed also found within this direction.

# Average intensity between sites at the various distances

Sites located within the OISST datasets shows no difference in duration of the events at different distances for each of the sites from the coastline. However, there is a difference in the mean intensity of events between sites (F=5.74, SS=5.2, p<0.0006). It is also evident that there is a significant difference in the mean intensity of the event at different distances from the coastline (F=260.17, SS=78.6, p<0.001), as well as differences in the duration in events (F=51.39, SS=15.53, p<0.001) and between sites . Similarly, the second dataset, the CMC dataset, results obtained from a  three-way ANOVA showed that there  was no a significant difference in the duration of events at different distances from the coastline for each of the sites (F=1.20, SS=0.41, p>0.05). However, there is a difference in the mean intensity of events at diferent distances (F=471.63, SS=162.0, p<0.05), as well as a signifcant difference in the duration of events (F=38.36, SS=13.2, p<0.006), and between sites (F=17.64, SS=18.2, p<0.002). It is also seen that there is no significant difference between the duration of upwelling between the various sites (F=0.22 SS=0.2, p<0.87). When looking at the in situ collected dataset, data collected at the coastline, we find no difference in the duration of events at the different sites (F=2.05, SS=1.84, p>0.11), however, we see a difference in the mean intensity of events at the different sites (F=2.05, SS=0.61, p>0.44).

# Plotting images

####TO do
- Plots
- Tables

# Discussion
